---
title: "Homework02"
author: "Yuchen Shao"
format: pdf
editor: visual
---

Link to GitHub repository：

<https://github.com/LAREINA-SHAO/STATS-506.git>

# Problem 1 - Dice Game

## a. Version 1

```{r}
#' Title Dice Game Function - Using a loop
#'
#' @param n The number of dice to roll
#'
#' @return  Total winnings
play_dice1 <- function(n){
  if (n < 0) {
    stop("Please input a non-negative number.")
  }
  
  # if no dice are rolled, return 0
  if (n == 0){
    return(0)
  }
  
  total_winnings <- 0
  # roll the dice n times
  die <- sample(1:6, n, replace = TRUE)

  for (i in 1:n) {
    # it costs $2 per game
    total_winnings <- total_winnings - 2 
    # on a roll of 3 or 5, win twice the roll
    if (die[i] == 3 | die[i] == 5) {
      total_winnings <- total_winnings + 2 * die[i]
    }
  }
  return(total_winnings)
}
```

## a. Version 2

```{r}
#' Title Dice Game Function - using built-in R vectorized functions
#'
#' @param n The number of dice to roll
#'
#' @return Total winnings
play_dice2 <- function(n){
  if (n < 0) {
    stop("Please input a non-negative number.")
  }
  
  # if no dice are rolled, return 0
  if (n == 0){
    return(0)
  }
  
  # roll the dice n times
  die <- sample(1:6, n, replace = TRUE)
  
  #winnings for rolls of 3 or 5
  winnings <- ifelse(die == 3 | die == 5, 2 * die, 0)
  total_winnings <- + sum(winnings) - 2 * n
  return(total_winnings)
}
```

## a. Version 3

```{r}
#' Title Dice Game Function - using table()
#'
#' @param n The number of dice to roll
#'
#' @return Total winnings
play_dice3 <- function(n){
  if (n < 0) {
    stop("Please input a non-negative number.")
  }
  
  # if no dice are rolled, return 0
  if (n == 0){
    return(0)
  }
  
  # roll the dice n times
  die <- sample(1:6, n, replace = TRUE)
  
  # create a frequency table (include 0 counts)
  die_table <- table(factor(die, levels = 1:6))
  
  # add together winnings and subtract out the total cost
  winnings_3 <- die_table["3"] * 2 * 3
  winnings_5 <- die_table["5"] * 2 * 5
  total_winnings <- winnings_3 + winnings_5 - 2 * n
  names(total_winnings) <- NULL
  return(total_winnings)
}
```

## a. Version 4

```{r}
#' Title Dice Game Function - using sapply
#'
#' @param n The number of dice to roll
#'
#' @return Total winnings
play_dice4 <- function(n){
  if (n < 0) {
    stop("Please input a non-negative number.")
  }
  
  # if no dice are rolled, return 0
  if (n == 0){
    return(0)
  }
  
  # roll the dice n times
  die <- sample(1:6, n, replace = TRUE)
  
  # apply the game rules using sapply
  winnings <- sapply(die, function(x) {
    if (x == 3 || x == 5) {
      return(2 * x)  
    } else {
      return(0)  
    }
  })
  total_winnings <- sum(winnings) - 2 * n
  return(total_winnings)
}  
```

## b.

```{r}
# total winnings for Version 1 with inputs of 3 and 3,000 dice rolls
c(play_dice1(3), play_dice1(3000))

# total winnings for Version 2 with inputs of 3 and 3,000 dice rolls
c(play_dice2(3), play_dice2(3000))

# total winnings for Version 3 with inputs of 3 and 3,000 dice rolls
c(play_dice3(3), play_dice3(3000))

# total winnings for Version 4 with inputs of 3 and 3,000 dice rolls
c(play_dice4(3), play_dice4(3000))
```

## c.

```{r}
# inputs 3
# set the seed for reproducibility
set.seed(111)  
result1 <- play_dice1(3)

# reset the seed to ensure same random sequence
set.seed(111)  
result2 <- play_dice2(3)

set.seed(111) 
result3 <- play_dice3(3)

set.seed(111) 
result4 <- play_dice4(3)

c(result1, result2, result3, result4)
```

```{r}
# inputs 3000
set.seed(111)  
result1 <- play_dice1(3000)

set.seed(111)  
result2 <- play_dice2(3000)

set.seed(111) 
result3 <- play_dice3(3000)

set.seed(111) 
result4 <- play_dice4(3000)

c(result1, result2, result3, result4)
```

## d.

```{r}
#install.packages("microbenchmark")
library(microbenchmark)
```

```{r}
#' Title Benchmarks for Dice Game Implementations
#'
#' @param n The number of dice rolls
#'
#' @return A `microbenchmark` object showing the performance of each version
BenchMarks <- function(n) {
  microbenchmark(
    play_dice1 = play_dice1(n),
    play_dice2 = play_dice2(n),
    play_dice3 = play_dice3(n),
    play_dice4 = play_dice4(n),
    times = 10  
  )
}

set.seed(111)

print(BenchMarks(1000))
print(BenchMarks(100000))
```

The function implemented using R's built-in vectorized functions is the fastest, while the function using sapply() performs the slowest. The function using a loop performs better than the one using table() when handling large inputs, but this advantage is negligible with smaller input sizes.

## e.

```{r}
#simulate 10000 times and pre-allocate a vector to store results
results <- numeric(10000)  

for (i in 1:10000) {
  # play 1000 dice rolls in each simulation
  results[i] <- play_dice2(1000)  
}

# Plot a histogram of the total winnings
hist(results, breaks = 50, main = "Monte Carlo Simulation of Total Winnings",
     xlab = "Total Winnings", col = "lightblue", border = "black")
abline(v = mean(results), col = "red", lwd = 2)

cat("Mean of total winnings:", mean(results), "\n")
```

This is not a fair game. The mean of total winnings across 10,000 simulations is significantly above zero, indicating that the player, on average, makes a profit.

# Problem 2 - Linear Regression

## a.

```{r}
cars <- read.csv("cars.csv")
names(cars)
```

```{r}
# rename the columns
names(cars) <- c("height", "length", "width", "driveline", "engine_type",
                 "hybrid","forward_gears", "transmission", "city_mpg",  
                 "fuel_type","highway_mpg", "classification", "ID", "make", 
                 "model_year", "year", "horsepower", "torque")
```

## b.

```{r}
gasoline_cars <- cars[cars$fuel == "Gasoline", ]
```

## c.

```{r}
# the distribution of highway_mpg
hist(cars$highway_mpg, main = "Distribution of Highway MPG", 
     xlab = "Highway MPG", col = "lightblue", border = "black")

summary(gasoline_cars$highway_mpg)
```

```{r}
# log transformation
gasoline_cars$log_highway_mpg <- log(gasoline_cars$highway_mpg)

# the distribution of log_highway_mpg 
hist(gasoline_cars$log_highway_mpg, main = "Log-Transformed Highway MPG",
     xlab = "Log(Highway MPG)", col = "purple", border = "black")
  
summary(gasoline_cars$log_highway_mpg)
```

The distribution of highway_mpg is highly right-skewed, suggesting the presence of outliers and a long tail of higher values. To normalize the distribution, I applied a logarithmic transformation, which helped reduce skewness and made the data more suitable for analysis.

##d.

```{r}
# convert 'year' to a factor (categorical variable)
gasoline_cars$year <- as.factor(gasoline_cars$year)
# fit a linear regression model
model <- lm(log_highway_mpg ~ torque + horsepower + height + length + width +
              year, data = gasoline_cars)
summary(model)
```

The estimated coefficient for torque is approximately -0.00229, meaning that for each additional unit of torque, the logarithm of highway MPG decreases by about 0.00229, holding all other variables constant. This indicates an inverse relationship between torque and highway fuel efficiency, where higher torque is associated with lower highway MPG. Additionally, this coefficient is highly significant, with a p-value less than 2e-16, confirming the robustness of the relationship.

## e.

```{r}
#install.packages("interactions")
library(interactions)
```

```{r}
hist(gasoline_cars$horsepower)
hist(gasoline_cars$torque)
```

```{r}
model_interaction  <- lm(log_highway_mpg ~ horsepower*torque + height + length +
                           width + as.factor(year), data = gasoline_cars)

interact_plot(model_interaction, pred = "torque", modx = "horsepower", 
                 # three different reasonable values of horsepower
                 modx.values = c(150, 200, 250), 
                 at = list(year = 2010),         
                 main.title = "Interaction between Torque and Horsepower",
                 xlab = "Torque", 
                 ylab = "Log Highway MPG",
                 legend.main = "Horsepower Levels")
```

The plot shows a negative relationship between torque and the logarithm of highway MPG, where increasing torque leads to a decrease in log_highway_mpg. The effect of torque on MPG is more pronounced for lower horsepower values, as shown by the steeper slopes for lower horsepower levels (150 and 200) compared to higher horsepower (250).

## f.

```{r}
# design matrix X for the linear regression
X <- model.matrix(log_highway_mpg ~ torque + horsepower + height + length +
                    width + as.factor(year), data = gasoline_cars)
y <- gasoline_cars$log_highway_mpg


# manually compute the OLS regression coefficients using matrix operations
# beta_hat = (X^T X)^(-1) X^T y
XtX_inv <- solve(t(X) %*% X)
XtY <- t(X) %*% y
beta_hat <- XtX_inv %*% XtY

cbind(model$coef, beta_hat)
```

The beta_hat calculated manually is the same as lm did prior.
