---
title: "506PS6"
author: "Yuchen Shao"
format: pdf
editor: visual
---

Link to GitHub repository：

<https://github.com/LAREINA-SHAO/STATS-506.git>

# Problem - Stratified Bootstrapping

```{r}
library(tidyverse)
library(dplyr)
library(DBI)     # For interfacing with a database

# Import the SQLite database of the Lahman data
lahman <- dbConnect(RSQLite::SQLite(), "lahman_1871-2022.sqlite")
dbListTables(lahman)
```
```{r}
dbListFields(lahman, "Fielding")
Fielding <- dbGetQuery(lahman, "SELECT * FROM Fielding")
dbDisconnect(lahman)
head(Fielding)
```
## a.
```{r}
fielding_data <- Fielding %>%
  filter(!is.na(PO), !is.na(A), !is.na(InnOuts), InnOuts > 0) %>%
  mutate(RF = 3*(PO + A) / InnOuts)

fielding_data %>%
  group_by(teamID) %>%
  summarise(avg_RF = mean(RF, na.rm = TRUE)) %>% print
```
### Without any parallel processing
```{r}
# Stratified Bootstrap Function
stratified_bootstrap <- function(data, target_col, strata_col, n_bootstrap) {
  strata_levels <- unique(data[[strata_col]]) #team
  bootstrap_samples <- list()
  
  # Perform stratified resampling
  for (level in strata_levels) {
    stratum_data <- data[data[[strata_col]] == level, ]
    if (nrow(stratum_data) == 0) next  # Skip if no data in the stratum
    sampled_data <- stratum_data[[target_col]][sample(1:nrow(stratum_data),
                                                      n_bootstrap, replace = TRUE)]
    bootstrap_samples[[level]] <- data.frame(
      bootstrap_sample = sampled_data,
      stratum = level
    )
  }
  
  # Combine results into a single data frame
  bootstrap_result <- do.call(rbind, bootstrap_samples)
  rownames(bootstrap_result) <- NULL
  return(bootstrap_result)
}

# Run the stratified bootstrap and calculate the execution_time
set.seed(123) 
n_bootstrap <- 1000
start_time <- Sys.time()
bootstrap_results1 <- stratified_bootstrap(fielding_data, "RF", "teamID", n_bootstrap)
end_time <- Sys.time()
execution_time1 <- end_time - start_time

# Estimate standard deviation for each team
bootstrap_sd1 <- bootstrap_results1 %>%
  group_by(stratum) %>%
  summarise(
    mean_RF = mean(bootstrap_sample, na.rm = TRUE),
    std_error_RF = sd(bootstrap_sample, na.rm = TRUE)
  )

# Display results
print(bootstrap_sd1)
```
### Using parallel processing with the parallel package
```{r}
library(parallel)

# Stratified Bootstrap Function with Parallel Processing
stratified_bootstrap_parallel <- function(data, target_col, strata_col, n_bootstrap) {
  strata_levels <- unique(data[[strata_col]])
  # Detect number of cores for parallel processing
  num_cores <- max(1, floor(detectCores() / 2)) 
  cl <- makeCluster(num_cores) # Create a cluster
  
  # Export required objects to the cluster
  clusterExport(cl, varlist = c("data", "target_col", "strata_col", "n_bootstrap"),
                envir = environment())

  # Parallelized bootstrap for each stratum
  bootstrap_samples <- parLapply(cl, strata_levels, function(level) {
    stratum_data <- data[data[[strata_col]] == level, ]
    if (nrow(stratum_data) == 0) return(NULL) # Skip if no data in stratum
    sampled_data <- stratum_data[[target_col]][sample(1:nrow(stratum_data),
                                                      n_bootstrap, replace = TRUE)]
    data.frame(
      bootstrap_sample = sampled_data,
      stratum = level
    )
  })
  
  stopCluster(cl) # Stop the cluster
 
  bootstrap_result <- do.call(rbind, bootstrap_samples)
  rownames(bootstrap_result) <- NULL
  return(bootstrap_result)
}

# Running the Bootstrap
set.seed(123) 
n_bootstrap <- 1000
start_time <- Sys.time()
bootstrap_results2 <- stratified_bootstrap_parallel(fielding_data, "RF", "teamID",
                                                    n_bootstrap)
end_time <- Sys.time()
execution_time2 <- end_time - start_time

# Summarizing the Results
bootstrap_sd2 <- bootstrap_results2 %>%
  group_by(stratum) %>%
  summarise(
    mean_RF = mean(bootstrap_sample, na.rm = TRUE),
    std_error_RF = sd(bootstrap_sample, na.rm = TRUE)
  )

print(bootstrap_sd2)
```
### Using futures with the future package
```{r, warning=FALSE}
library(furrr)
# Stratified Bootstrap Function with Future Package
stratified_bootstrap_future <- function(data, target_col, strata_col, n_bootstrap) {
  strata_levels <- unique(data[[strata_col]])
 
  plan(multisession)
  
  # Perform bootstrap sampling 
  bootstrap_samples <- future_map(strata_levels, function(level) {
    stratum_data <- data[data[[strata_col]] == level, ]
    if (nrow(stratum_data) == 0) return(NULL) # Skip if no data in stratum
    sampled_data <- stratum_data[[target_col]][sample(1:nrow(stratum_data),
                                                      n_bootstrap, replace = TRUE)]
    data.frame(
      bootstrap_sample = sampled_data,
      stratum = level
    )
  })
  
  bootstrap_result <- do.call(rbind, bootstrap_samples)
  rownames(bootstrap_result) <- NULL
  plan(sequential)  # Reset to sequential
  return(bootstrap_result)
}

# Running the Bootstrap
set.seed(123) 
n_bootstrap <- 1000
start_time <- Sys.time()
bootstrap_results3 <- stratified_bootstrap_future(fielding_data, "RF", "teamID",
                                                  n_bootstrap)
end_time <- Sys.time()
execution_time3 <- end_time - start_time

# Summarizing the Results
bootstrap_sd3 <- bootstrap_results3 %>%
  group_by(stratum) %>%
  summarise(
    mean_RF = mean(bootstrap_sample, na.rm = TRUE),
    std_error_RF = sd(bootstrap_sample, na.rm = TRUE)
  )

print(bootstrap_sd3)
```
## b.
```{r}
bootstrap_sd1$method <- "No Parallel"
bootstrap_sd2$method <- "Parallel"
bootstrap_sd3$method <- "Future"

top_10_results <- bind_rows(bootstrap_sd1, bootstrap_sd2, bootstrap_sd3) %>%
  group_by(method) %>%
  arrange(desc(mean_RF)) %>%
  slice_head(n = 10) %>%
  mutate(mean_RF = round(mean_RF, 3),
         std_error_RF = round(std_error_RF, 3))

cat("Top 10 Results - No Parallel\n")
print(top_10_results[top_10_results$method == 'No Parallel', ])

cat("\nTop 10 Results - Parallel\n")
print(top_10_results[top_10_results$method == 'Parallel', ])

cat("\nTop 10 Results - Future\n")
print(top_10_results[top_10_results$method == 'Future', ])
```
## c.
```{r}
Execution_time_difference <- data.frame(
  Method = c("No Parallel", "Parallel", "Future"),
  Execution_Time = c(execution_time1, execution_time2, execution_time3)
)
print(Execution_time_difference)
ggplot(Execution_time_difference, aes(x = Method, y = Execution_Time, fill = Method)) +
  geom_bar(stat = "identity") +
  labs(title = "Execution Time Comparison",
       x = "Method", y = "Execution Time (seconds)") +
  theme_minimal()
```
```{r}
# Extract unique teams for each method
no_parallel_teams <- top_10_results %>%
  filter(method == "No Parallel") %>%
  pull(stratum)

parallel_teams <- top_10_results %>%
  filter(method == "Parallel") %>%
  pull(stratum)

future_teams <- top_10_results %>%
  filter(method == "Future") %>%
  pull(stratum)

# Find the intersection of teams
common_teams <- Reduce(intersect, list(no_parallel_teams, parallel_teams, future_teams))

cat("Teams in the Top 10 for All Methods:\n")
print(common_teams)
```


```{r}
# Stability analysis: Variance in mean_RF estimates
stability_test <- function(results1, results2, results3) {
  combined <- bind_rows(
    results1 %>% mutate(method = "No Parallel"),
    results2 %>% mutate(method = "Parallel"),
    results3 %>% mutate(method = "Future")
  )
  
  combined %>%
    group_by(method) %>%
    summarise(
      Variance_Mean_RF = var(mean_RF, na.rm = TRUE),
      Variance_Std_Error_RF = var(std_error_RF, na.rm = TRUE)
    )
}

# Run stability analysis
stability_results <- stability_test(bootstrap_sd1, bootstrap_sd2, bootstrap_sd3)
print(stability_results)
```

The No Parallel method is the fastest due to its simplicity and lack of parallelization overhead. The Parallel method, at around 1.5 seconds, is slower due to the cost of setting up clusters and managing multiple cores, but it balances performance and resource utilization well. The Future method is the slowest, possibly due to higher overhead in managing futures. The calculated average RF (mean_RF) for each team varies across the different methods, but the differences are not substantial and remain within a consistent range. Notably, teams such as "LS1," "RC1," "ELI," "MLU," "KEO," "BLA," and "RIC" consistently rank among the teams with the 10 highest RF across all three approaches, demonstrating their stability in performance regardless of the method used.







